Specifgic Gene or Metabolite Levels

### 2025-01-13 GeneExpressionBoxPlots.Rmd
data files

---
title: "rac1 gene expression"
author: "John Santiago"
date: "2024-05-24"
output: html_document
---

@@

```{r include = F, echo = F}

library(plotly)
library(edgeR)
library(heatmaply)
library(org.Dm.eg.db)
library(KEGGREST)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

groups = groups[colnames(cpmdata),]

@@

nodeID = as.list(org.Dm.egENZYME)

##subsetting genes for patterns

##relish FBgn0014018
##attacin FBgn0041579
##cecropin FBgn0000276
##diptericin FBgn0004240
##drosocin FBgn0010388
##defensin FBgn0010385
##drosomycin FBgn0283461
##metchnikowin FBgn0014865
##sod1 FBgn0003462
##sod2 FBgn0010213
##catalase FBgn0000261
##gstd1 FBgn0001149
## arc1 FBgn0033926
##rac1
##cgs = c('CG10592','CG10827','CG1462','CG16771','CG1809','CG3264','CG3290','CG3292','CG5150','CG5361','CG5656','CG8105','CG8147')
##mle 'FBgn0002774'
##gstD FBgn0001149
##Tudor-SN FBgn0035121
##atf4 FBgn0000370
##thor FBgn0261560
##sdhbl FBgn0030975
##ND-51L FBgn0034251
##kayak (AP-1) FBgn0001297
## Jra (AP-1) FBgn0001291
#sad FBgn0003312

##Specific enzyme
enzyme.ens = nodeID[grep('3.1.3.5',nodeID)]
enzyme.genes = GeneIDKey[GeneIDKey$ensembl %in% names(enzyme.ens)[c(1,3:5,7)],]

> GeneIDKey[intersect(s1,intersect(s2,intersect(s3,s4))),'Symbol']
 [1] "Cyp4p1"  "CG13962" "Cpr51A"  "CG11438" "Sgs5bis" "sad"    
 [7] "Lip2"    "CG4757"  "CG14229" "CG15406" "CG3270"  "CG9877" 
[13] "CG13905" "Zip71B"  "CG17027" "nom"     "CG16727" "CG9702" 
[19] "CG43740"
> GeneIDKey[intersect(s1,intersect(s2,intersect(s3,s4))),'FBgn']
 [1] "FBgn0015037" "FBgn0032824" "FBgn0033942" "FBgn0037164"
 [5] "FBgn0038523" "FBgn0003312" "FBgn0024740" "FBgn0027584"
 [9] "FBgn0031059" "FBgn0031517" "FBgn0033093" "FBgn0034819"
[13] "FBgn0035176" "FBgn0036461" "FBgn0036553" "FBgn0037617"
[17] "FBgn0038719" "FBgn0039787" "FBgn0263997"

FBgn = 'FBgn0032513'
name = 'STARD7'
FBgn = 'FBgn0028932'
name = 'CG16890'

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), condition = groups$Group, group = paste0(groups$Genotype, groups$Sex))

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), 
                  condition = groups$Group, 
                  group = paste0(groups$Genotype, groups$Sex),
                  TKT = groups$TKT)

A4V.groups = as.vector(colnames(A4V.cpm))

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]), 
                  condition = A4V.groups, 
                  group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)

```{r}

```{r}

@@

metab = 'D-sedoheptulose-7-phosphate'
name = 'D-sedoheptulose-7-phosphate'

@@

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(metab$data), max(metab$data)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

@@

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

@@

>>>>

---
title: "Specifgic Gene or Metabolite Levels"
author: "John Santiago"
date: "2024-05-24"
output: html_document
---

@@

```{r include = F, echo = F}
##Load Libraries and Data

library(plotly)
library(edgeR)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"

TKT.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)

TKT.groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

TKT.groups = TKT.groups[colnames(TKT.cpm),]

@@

@@

```{r echo = F}
##Function to plot all G85R groups

plot.G85R = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

gene = data.frame(cpm = as.numeric(TKT.cpm[FBgn,]), condition = TKT.groups$Group, group = paste0(TKT.groups$Genotype, TKT.groups$Sex))

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)
}

if(nrow(ID.row) != 1){
  ID.row
}
}

```{r echo = F}
##Function to plot a specific gene expression levels in all A4V groups

plot.A4V = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
  FBgn = ID.row$FBgn
  name = ID.row$Symbol
  A4V.groups = as.vector(colnames(A4V.cpm))

  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]),
                    condition = A4V.groups,
                    group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, start = nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

  boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
  points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
  legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)
}

if(nrow(ID.row) != 1){
  ID.row
}
}

```{r echo = F}
##Plot levels of a specific metabolite

plot.metab = function(metabID){

ID.row = row.names(norm.data)[grep(ID, row.names(norm.data))]

if(length(ID.row) == 1){
metab = ID.row
name = ID.row

@@

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(na.omit(metab$data)), max(na.omit(metab$data))),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

@@

}

if(length(ID.row) != 1){
  ID.row
}

}

@@

ID = "0000018"
plot.G85R(ID)

```{r echo = F}
##Tailared code to plot a specific gene expression levels in all A4V groups

ID = "0000018"

plot.A4V(ID)
```

metabID = 'NADP'

plot.metab(metabID)

```{r echo = F}
##Tailared code to plot a subset of groups in the TKT dataset

##ID = "0000018"

ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

gene = data.frame(cpm = as.numeric(TKT.cpm[FBgn,]), 
                  condition = TKT.groups$Group, 
                  group = paste0(TKT.groups$Genotype, TKT.groups$Sex),
                  TKT = TKT.groups$TKT)

}

if(nrow(ID.row) != 1){
  ID.row
}

```{r echo = F}
##Plot expression levels for a specific gene in the gbbOE dataset

##ID = "0000018"

ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

}

if(nrow(ID.row) != 1){
  ID.row
}
```

```{r echo = F}
##Plot expression levels for a specific gene in the gbb-null dataset

##ID = "0000018"

ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

}

if(nrow(ID.row) != 1){
  ID.row
}
```
